Analyzing DECREASE trials for extent of data fabrication

### 2017-03-09 submission/manuscript.tex
Revisions manuscript, draft discussion.

\setlength{\droptitle}{-2em}
  \title{Analyzing DECREASE trials for extent of data fabrication}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{02 March, 2017}

@@

The trials subject to research misconduct were the Dutch DECREASE-I and
DECREASE-IV trials\textsuperscript{2--4}. The committees that
investigated the integrity of the DECREASE trials reported that data
fabrication was likely but that the extent of the data fabrication
remained unclear\textsuperscript{2--4}. Moreover, the latest guidelines
still recommend the usage of beta-blockers in the perioperative period
in certain cases\textsuperscript{8,9}. Considering these consequences,
we aim to estimate the extent of data fabrication in the DECREASE
studies\textsuperscript{10,11} to further the debate on using
beta-blockers in the perioperative period of CAD patients undergoing
non-cardiac surgery.

The reports on the integrity of the DECREASE trials primarily focused on
the provenance of the raw data but did not investigate the extent to
which the DECREASE trials deviated from comparable trials. The
provenance is primarily concerned with the origins of the data,
verifying things such as (but not limited to) the informed consent and
whether data corresponded to patient files. The committee reports did
not neglect statistical evaluation however: a statistical expert
evaluated the applicability of forensic statistical
methods\textsuperscript{4} to evaluate results of trials separately
(i.e., DECREASE-I, DECREASE-IV). Nonetheless, comparing across trials is
a method that has previously been used to monitor trial data quality or
to test for potential data anomalies\textsuperscript{12}. Moreover, this
method has previously proven to be effective in detecting data
fabrication\textsuperscript{13}. Evaluating the DECREASE trials to other
published trials studying the effectiveness of beta-blockers with
respect to perioperative mortality could prove informative of the
potential extent of the fabrication in the DECREASE trials.

However, the effectiveness of perioperative beta-blockade is obfuscated
by both trials afflicted by data fabrication and the type of
beta-blocker used in the trial. Various kinds of beta-blockers could
vary in their pharmalogical effectiveness with respect to perioperative
mortality, and anomalies in the effectiveness of the DECREASE trials
might be due to differences in the beta-blockers used {[}@{]} and not
purely due to data fabrication (given not all data points can be
considered fabricated at this point).

To statistically investigate the evidence of data fabrication in the
DECREASE studies\textsuperscript{10,11}, we took three steps. First, we
reproduced the findings from a 2014 meta-analysis\textsuperscript{7}
that contained sufficient information to estimate the deviation of the
DECREASE trials from other published trials on beta-blockers. We also
include type of beta-blocker to inspect whether this is predictive of
the effect of beta-blockers on perioperative mortality. Second, we
evaluated the the probability that the DECREASE trial results occurred
assuming no data fabrication. Third, we reversed this assumption and
assumed that data fabrication did occur and estimated how many data
points would have to be fabricated to reproduce the results of the
DECREASE trials, if the other published trials are regarded as
estimating the true effect of beta-blockers on perioperative mortality
in CAD patients undergoing non-cardiac surgery.

@@

Second, we meta-analyzed all studies combined, including a dummy
predictor for the DECREASE and non-DECREASE studies to reproduce results
presented in Figure 4 of the 2014 meta-analysis\textsuperscript{7}.
Surprisingly, our results showed stronger evidence against equal
subgroups than the original meta-analysis\textsuperscript{7} (original:
\(\chi^2(1)=3.91,p=.05\), reproduced \(\chi^2(1)=6.12,p=0.013\)).
Additionally, the original analyses showed substantial residual
heterogeneity (\(I^2=74.4\)\%) whereas we found no residual heterogeneit
(\$I\^{}2=\$0\%). Different variance estimates (e.g.,
\texttt{DerSimonian-Laird} instead of \texttt{REML}) did not resolve
this difference. We tried to clarify these discrepancies by e-mailing
the original authors, but did not receive a response. Nonetheless, the
broad strokes of the meta-regression confirmed that the DECREASE trials
were the determining predictor for the effectiveness of beta-blockers
(including DECREASE: \(RR=0.496\); excluding DECREASE: \(RR=1.28\)).

@@

Based on the non-DECREASE estimates from Step 1, we estimated the
probability of obtaining the results in the DECREASE trials. To this
end, we assumed that the non-DECREASE trials provide a valid
representation of the true effect of beta-blockers on perioperative
mortality (similar to Bouri et al.\textsuperscript{7}; but see the
discussion). The estimated probability is also known as the veracity of
the data\textsuperscript{17}, which indicates the probability of the
observed data under a given true effect. We assumed that the
non-DECREASE studies estimated the true effect distribution of
perioperative beta-blockade on mortality, not perturbed by publication
bias due to statistical (non)significance. Publication bias was assumed
to not be a problem because a substantial number of nonsignificant
effects are included in the dataset (9 of 11 nonsignificant results).

@@

Based on the estimated mean log RR and its credible interval in the
non-DECREASE studies, we computed the probability of the observed log RR
in the DECREASE trials. The estimates of the non-DECREASE studies were
obtained from Step 1, which include the estimated log RR (i.e., 0.25),
and its 95\% credibility interval as provided by the package
\texttt{metafor} (i.e., {[}0.01; 0.484{]}. The meta-analysis model
assumes a normal distribution of population effects with the estimated
effect as the mean of the distribution. The 95\% credibility interval
denotes the bounds of the normal distribution that covers 95\% of the
density, where the standard deviation is calculated as the distance from
the mean to either bound, divided by 1.96. This allows for an
approximation of the population effect distribution, as depicted in
Figure 1.

@@

Based on the estimated effect distribution from the non-DECREASE trials,
we calculated the probability of each DECREASE trial result, or a more
extreme result. In other words, we computed the \emph{p}-value for the
hypothesis that the DECREASE studies occurred, assuming the information
available from the other trials is informative of the true population
effect.

@@

Figure 1 indicates that the DECREASE trials are highly unlikely under
the estimated effect distribution based on the non-DECREASE trials. More
specifically, the results from DECREASE-I (or more extreme) have a
probability of \(0\) (less than 1 in a quintillion) and the results from
DECREASE-IV have a probability of \(0\) (3 in a billion). This indicates
the DECREASE trial results are unlikely to have come from the same
population effect distribution as the non-DECREASE trials. Moreover,
observing two of such extremely unlikely results jointly, as in the
DECREASE trials, is highly improbable, \(0\). This would therefore
indicate that the DECREASE trials are severely different from the
non-DECREASE trials.

Results from Step 1 indicated that no between-study variance (i.e.,
homogeneity; \(\tau^2=0\)) of the effects was observed; given the small
number of studies included (i.e., 9) this estimate is uncertain,
however. We conducted sensitivity analyses to see how dependent results
are on the heterogeneity estimate. The probability of observing the
DECREASE trials stays approximately below 1 out of 1000 until the
variance estimate is 0.25; the probability stays approximately below 1
out of 100 until the variance estimate is 0.43 (see Figure 2). To put
these numbers into context, a variance of 0.25 would suggest that
results of perioperative beta-blockade vary substantially even if
perioperative beta-blockade has no effect whatsoever (RRs between 0.779
and 1.284 in 64\% of the cases) all due to contextual circumstances of
the study.

@@

For DECREASE-I (\$N=112), the 95\% confidence interval for the estimated
number of fabricated data points is {[}0 - 112{]} or {[}2 - 112{]} when
based on a point estimate or a more uncertain distribution estimate,
respectively. The left column of Figure 3 depicts the \(p_F\) per
\emph{X} fabricated data points (top panel) and the bounds of the
confidence interval when the degree of confidence is altered (lower
panel). Staying clearly between the dotted lines in the top panel,
depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent
that the degree of uncertainty is too high to make any reasonable
estimates about the number of fabricated data points with sufficient
confidence. This is partly due to the small sample size of the
DECREASE-I trial (i.e., \(N=112\)) and the availability of just the
summary results. Only when the degree of confidence is lowered to around
75\% does the interval not span the entire sample size. As such, based
on the summary results, little can be said about the extent of the data
fabrication that occurred in the DECREASE-I trial, affirming the
conclusions of the original committee report\textsuperscript{4}.

For DECREASE-IV (\$N=1066), the 95\% confidence interval for the
estimated number of fabricated data points is {[}0 - 1053{]} or {[}8 -
1066{]} when based on a point estimate or a more uncertain distribution
estimate, respectively. The relatively minor difference between the
estimates indicates that there is a high degree of confidence that data
fabrication did occur based on the difference of the trial results
alone. Nonetheless, the range of potentially fabricated data points is
still estimated at approximately 1000; this indicates that the summary
results are insufficient to provide more than an estimated lowerbound.
This indicates that it is possible not all data were fabricated (i.e.,
\(N=1066\)), increasing the importance of well-documented data
provenance to discern between genuine and falsified data. This affirms
the conclusion of the scientific integrity committee that the results of
the DECREASE-IV trial are ``scientifically incorrect''
(p.11)\textsuperscript{3}.

The results also highlight that, despite the lack of availability of the
raw data, summary results from larger samples allow estimation of the
number of fabricated data points when similar trials are available.
Moreover, larger trials result in relatively more certainty (e.g.,
DECREASE-IV) about the estimated number of fabricated data points using
the inversion method compared to smaller trials (e.g., DECREASE-I),
although much residual uncertainty remains. This increased certainty is
due to decreased standard errors at the population level of the
estimated effects, resulting in higher sensitivity to data anomalies.
Nonetheless, substantially less information is available in just the
summary results, and raw data availability would improve results
verification and the detection of potential anomalies. These results
also highlight that in order to prevent detection, it would be in the
fabricators' interest to fabricate small studies even if raw data are
hidden away (assuming the fabricator wants to remain undetected).

@@

Throughout steps 1, 2, and 3 we provided evidence that the , the
DECREASE trials are highly unlikely to come from a genuine data pool,
and that there is a high degree of confidence that data was fab

We note that the variance in the non-DECREASE studies was estimated at
0, indicating that there would supposedly be a fixed-effect of
beta-blockers on perioperative mortality despite differences in type of
beta-blockers administered and dosage/duration of the beta-blockers.
This homogeneity is not excessive, considering that the
\emph{Q}-statistic is sufficiently large to begin with (i.e.,
5).\textsuperscript{19} Nonetheless, the original non-DECREASE trials
show substantive variation in the treatment, for example the usage of
different beta-blockers (i.e., atenolol, propanolol, bisoprolol,
metoprolol) in different (standardized) dosages, which have been argued
to affect effectiveness of the interventions\textsuperscript{20}. As
such, the lack of heterogeneity can be genuine and indicative of no
difference between these different beta-blockers and their
dosage/duration. Nonetheless, it could also be due to uncertainty in the
estimate of heterogeneity \(\tau^2\) due to the small set of studies
included.

However, due to various types of beta-blockers, evidence might be
moderated\textsuperscript{20}.

Additionally, during the process of this research paper, the first
author tried to ascertain a raw dataset of one of the DECREASE trials
via a Freedom of Information request (FOI). The Erasmus MC refused to
share these data

The ESC/ESA guidelines from 2014{[}@{]} were published in October 2014,
whereas the meta-analysis indicating a reversal of the effectiveness of
beta-blockers was published online at the end of July 2013{[}@{]}. The
committee revising the ESC/ESA guidelines stated: ``The respective
writing committees independently performed their literature review and
analysis, and then developed their
recommendations.''\textsuperscript{21} Regardless, the revisions made
``recommend continuation of beta-blocker therapy in the perioperative
period in patients currently receiving this medication'' and do not
recommend the ``initiation of beta-blockers in patients undergoing
low-risk surgery''\textsuperscript{22}. Review pieces note no
substantial clinical change with respect to
beta-blockers.\textsuperscript{23,24}

The revisions to the ESC/ESA guidelines do not discourage use of
beta-blockers, despite the obvious lack of evidence for systematic
evidence of the effectiveness of beta-blockers to reduce perioperative
mortality in trials. If anything, the trials show that insufficient
evidence has been collected. This becomes apparent when the estimated
effect for all non-DECREASE trials is inspected; its \emph{p}-value is
.04, which is insufficient evidence and is actually more likely when
there is no effect in of beta-blockers in reality\textsuperscript{25}

In order , the first author tried to ascertain the data of DECREASE VI,
the only DECREASE study for which a dataset is still
available\textsuperscript{1} via a Freedom of Information (FOI) request.
Within this paper, we inspected data anomalies based on summary results,
which contain less information to be able to determine what isThe board
of Erasmus Medical Centrum (EMC) was unwilling to comply with this
request, stating that it was not a matter of the board. Regardless of
whether it was a board related matter,

@@

\hypertarget{ref-Fleishere278}{}
9. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 acc/aha
guideline on perioperative cardiovascular evaluation and management of
patients undergoing noncardiac surgery. \emph{Circulation}.
2014;130(24):e278-e333.
doi:\href{https://doi.org/10.1161/CIR.0000000000000106}{10.1161/CIR.0000000000000106}.

\hypertarget{ref-dunkelgrun2009}{}
10. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
fluvastatin for the reduction of perioperative cardiac mortality and
myocardial infarction in intermediate-risk patients undergoing
noncardiovascular surgery: A randomized controlled trial (DECREASE-IV).
\emph{Annals of surgery}. 2009;249(6):921-926.
doi:\href{https://doi.org/10.1097/SLA.0b013e3181a77d00}{10.1097/SLA.0b013e3181a77d00}.

\hypertarget{ref-poldermans1999}{}
11. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
perioperative mortality and myocardial infarction in High-Risk patients
undergoing vascular surgery. \emph{The New England journal of medicine}.
1999;341(24):1789-1794.
doi:\href{https://doi.org/10.1056/NEJM199912093412402}{10.1056/NEJM199912093412402}.

\hypertarget{ref-Buyse1999-jq}{}
12. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
prevention, detection and treatment of fraud in clinical trials.
\emph{Statistics in medicine}. 1999;18(24):3435-3451.
doi:\href{https://doi.org/10.1002/(SICI)1097-0258(19991230)18:24\%3C3435::AID-SIM365\%3E3.0.CO;2-O}{10.1002/(SICI)1097-0258(19991230)18:24\textless{}3435::AID-SIM365\textgreater{}3.0.CO;2-O}.

\hypertarget{ref-Knepper2016-la}{}
13. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
clinical trials: Best practices for detecting data anomalies suggestive
of fabrication or misconduct. \emph{Therapeutic Innovation \& Regulatory
Science}. 4\textasciitilde{}feb 2016.
doi:\href{https://doi.org/10.1177/2168479016630576}{10.1177/2168479016630576}.

@@

\hypertarget{ref-ioannidis2006}{}
19. Ioannidis JPA, Trikalinos TA, Zintzaras E. Extreme between-study
homogeneity in meta-analyses could offer useful insights. \emph{Journal
of clinical epidemiology}. 2006;59(10):1023-1032.
doi:\href{https://doi.org/10.1016/j.jclinepi.2006.02.013}{10.1016/j.jclinepi.2006.02.013}.

\hypertarget{ref-klei2015}{}
20. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
{[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
Tijdschrift voor Geneeskunde}. 2015;159:A9798.

\hypertarget{ref-Kristensen_2014}{}
21. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
2014;31(10):517-573.
doi:\href{https://doi.org/10.1097/eja.0000000000000150}{10.1097/eja.0000000000000150}.

\hypertarget{ref-guarracino2015}{}
22. Guarracino F PH Baldassarri R. Revised esc/esa guidelines on
non-cardiac surgery: Cardiovascular assessment and management.
implications for preoperative clinical evaluation. \emph{Minerva
Anestesiology}. 2015;2:225-233.

\hypertarget{ref-Port2016}{}
23. Port SC. 2014 esc/esa guidelines on noncardiac surgery:
Cardiovascular assessment and management. \emph{Journal of Nuclear
Cardiology}. 2016:1-3.
doi:\href{https://doi.org/10.1007/s12350-016-0641-x}{10.1007/s12350-016-0641-x}.

\hypertarget{ref-Kristensen2016}{}
24. Kristensen SD. 2014 esc/esa guidelines on noncardiac surgery:
Cardiovascular assessment and management. \emph{Journal of Nuclear
Cardiology}. 2016:1-3.
doi:\href{https://doi.org/10.1007/s12350-016-0642-9}{10.1007/s12350-016-0642-9}.

\hypertarget{ref-van_Aert_2016}{}
25. Aert RCM van, Wicherts JM, Assen MALM van. Conducting meta-analyses
based on p values: Reservations and recommendations for applying
p-uniform and p-curve. \emph{Perspectives on Psychological Science}.
2016;11(5):713-729.
doi:\href{https://doi.org/10.1177/1745691616650874}{10.1177/1745691616650874}.

>>>>

\setlength{\droptitle}{-2em}
  \title{Analyzing DECREASE trials for extent of data fabrication}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{09 March, 2017}

@@

The trials subject to research misconduct were the Dutch DECREASE-I and
DECREASE-IV trials\textsuperscript{2--4}. The committees that
investigated the integrity of the DECREASE trials reported that data
fabrication was likely but that the extent of the data fabrication
remained unclear\textsuperscript{2--4}. Moreover, the latest guidelines
still recommend the usage of beta-blockers in the perioperative period
in certain cases\textsuperscript{{\textbf{???}},8}. Considering these
consequences, we aim to estimate the extent of data fabrication in the
DECREASE studies\textsuperscript{9,10} to further the debate on using
beta-blockers in the perioperative period of CAD patients undergoing
non-cardiac surgery.

The reports on the integrity of the DECREASE trials primarily focused on
the provenance of the raw data but did not investigate the extent to
which the DECREASE trials deviated from comparable trials. The
provenance is primarily concerned with the origins of the data,
verifying things such as (but not limited to) the informed consent and
whether data corresponded to patient files. The committee reports did
not neglect statistical evaluation however: a statistical expert
evaluated the applicability of forensic statistical
methods\textsuperscript{4} to evaluate results of trials separately
(i.e., DECREASE-I, DECREASE-IV). Nonetheless, comparing across trials is
a method that has previously been used to monitor trial data quality or
to test for potential data anomalies\textsuperscript{11}. Moreover, this
method has previously proven to be effective in detecting data
fabrication\textsuperscript{12}. Evaluating the DECREASE trials to other
published trials studying the effectiveness of beta-blockers with
respect to perioperative mortality could prove informative of the
potential extent of the fabrication in the DECREASE trials.

However, the effectiveness of perioperative beta-blockade is obfuscated
by both trials afflicted by data fabrication and the type of
beta-blocker used in the trial. Various kinds of beta-blockers could
vary in their pharmalogical effectiveness with respect to perioperative
mortality, and anomalies in the effectiveness of the DECREASE trials
might be due to differences in the beta-blockers
used\textsuperscript{13} and not purely due to data fabrication (given
not all data points can be considered fabricated at this point).

To statistically investigate the evidence of data fabrication in the
DECREASE studies\textsuperscript{9,10}, we took three steps. First, we
reproduced the findings from a 2014 meta-analysis\textsuperscript{7}
that contained sufficient information to estimate the deviation of the
DECREASE trials from other published trials on beta-blockers. We also
include type of beta-blocker to inspect whether this is predictive of
the effect of beta-blockers on perioperative mortality. Second, we
evaluated the probability that the DECREASE trials occurred assuming no
data fabrication. Third, we reversed this assumption and assumed that
data fabrication did occur and estimated how many data points would have
to be fabricated to reproduce the results of the DECREASE trials, if the
other published trials are regarded as estimating the true effect of
beta-blockers on perioperative mortality in CAD patients undergoing
non-cardiac surgery.

@@

Second, we meta-analyzed all studies combined, including a dummy
predictor for the DECREASE and non-DECREASE studies to reproduce results
presented in Figure 4 of the 2014 meta-analysis\textsuperscript{7}.
Surprisingly, our results showed stronger evidence against equal
subgroups than the original meta-analysis\textsuperscript{7} (original:
\(\chi^2(1)=3.91,p=.05\); reproduced: \(\chi^2(1)=6.12,p=0.013\)).
Additionally, the original analyses showed substantial residual
heterogeneity (\(I^2=74.4\)\%) whereas we found no residual
heterogeneity (\(I^2=0\)\%). Different variance estimates (e.g.,
\texttt{DerSimonian-Laird} instead of \texttt{REML}) did not resolve
this difference. We tried to clarify these discrepancies by e-mailing
the original authors, but did not receive a response. Nonetheless, the
broad strokes of the meta-regression confirmed that the DECREASE trials
were the determining predictor for the effectiveness of beta-blockers
(including DECREASE: \(RR=0.496\); excluding DECREASE: \(RR=1.28\)).

@@

Based on the non-DECREASE estimates from Step 1, we estimated the
probability of obtaining the results in the DECREASE trials. To this
end, we assumed that the non-DECREASE trials provide a valid
representation of the true effect of beta-blockers on perioperative
mortality (similar to Bouri et al.\textsuperscript{7}). The estimated
probability is also known as the veracity of the
data\textsuperscript{17}, which indicates the probability of the
observed data under a given true effect. We assumed that the
non-DECREASE studies estimated the true effect distribution of
perioperative beta-blockade on mortality, not perturbed by publication
bias due to statistical (non)significance. Publication bias was assumed
to not be a problem because a substantial number of nonsignificant
effects are included in the dataset (9 of 11 results are
nonsignificant).

@@

Based on the estimated mean log RR and its credible interval in the
non-DECREASE studies, we computed the probability of the observed log RR
in the DECREASE trials. The estimates of the non-DECREASE studies were
obtained from Step 1, which include the estimated log RR (i.e., 0.25),
and its 95\% credibility interval as provided by the package
\texttt{metafor} (i.e., {[}0.01; 0.484{]}). The meta-analysis model
assumes a normal distribution of population effects with the estimated
effect as the mean of the distribution. The 95\% credibility interval
denotes the bounds of the normal distribution that covers 95\% of the
density, where the standard deviation is calculated as the distance from
the mean to either bound, divided by 1.96. This allows for an
approximation of the population effect distribution, as depicted in
Figure 1.

@@

Based on the estimated effect distribution from the non-DECREASE trials,
we calculated the probability of each DECREASE trial result, or a more
extreme result. In other words, we computed the \emph{p}-value for the
null hypothesis that the DECREASE trials arise from the same effect
distribution as the non-DECREASE trials. This assumes that the
information available from the other trials is informative of the true
population effect.

@@

Figure 1 indicates that the DECREASE trials are highly unlikely under
the estimated effect distribution based on the non-DECREASE trials. More
specifically, the results from DECREASE-I (or more extreme) have a
probability of \(2.9610213\times 10^{-53}\) (less than 1 in a
quintillion) and the results from DECREASE-IV have a probability of
\(3.0974376\times 10^{-9}\) (3 in a billion). This indicates the
DECREASE trial results are unlikely to have come from the same
population effect distribution as the non-DECREASE trials. Moreover,
observing two of such extremely unlikely results jointly, as in the
DECREASE trials, is nearly impossible, \(9.1715786\times 10^{-62}\).
Hence, this result indicates that the DECREASE trials are severely
different from the non-DECREASE trials.

Results from Step 1 indicated that no between-trial variance (i.e.,
homogeneity; \(\tau^2=0\)) of the effects was observed; given the small
number of trials included (i.e., 9) this estimate is uncertain, however.
We conducted sensitivity analyses to see how dependent results are on
the heterogeneity estimate. The probability of observing the DECREASE
trials stays approximately below 1 out of 1000 until the variance
estimate is 0.25; the probability stays approximately below 1 out of 100
until the variance estimate is 0.43 (see Figure 2). To put these numbers
into context, a variance of 0.25 would suggest that results of
perioperative beta-blockade vary substantially even if perioperative
beta-blockade has no effect whatsoever (RRs between 0.779 and 1.284 in
64\% of the cases) all due to contextual circumstances of the study.

@@

For DECREASE-I (\(N=112\)), the 95\% confidence interval for the
estimated number of fabricated data points is {[}0 - 112{]} or {[}2 -
112{]} when based on a point estimate or a more uncertain distribution
estimate, respectively. The left column of Figure 3 depicts the \(p_F\)
per \emph{X} fabricated data points (top panel) and the bounds of the
confidence interval when the degree of confidence is altered (lower
panel). Staying clearly between the dotted lines in the top panel,
depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent
that the degree of uncertainty is too high to make any reasonable
estimates about the number of fabricated data points with sufficient
confidence. This is partly due to the small sample size of the
DECREASE-I trial (i.e., \(N=112\)) and the availability of just the
summary results. Only when the degree of confidence is lowered to around
75\% does the interval not span the entire sample size. As such, based
on the summary results, little can be said about the extent of the data
fabrication that occurred in the DECREASE-I trial, affirming the
conclusions of the original committee report\textsuperscript{4}.

For DECREASE-IV (\(N=1066\)), the 95\% confidence interval for the
estimated number of fabricated data points is {[}3 - 1059{]} or {[}8 -
1066{]} when based on a point estimate or a more uncertain distribution
estimate, respectively. The relatively minor difference between the
estimates indicates that there is a high degree of confidence that data
fabrication did occur based on the difference of the trial results
alone. Nonetheless, the range of potentially fabricated data points is
still estimated at approximately 1000; this indicates that the summary
results are insufficient to provide more than an estimated lowerbound.
This indicates that it is possible not all data were fabricated (i.e.,
\(N=1066\)), increasing the importance of well-documented data
provenance to discern between genuine and falsified data. This affirms
the conclusion of the scientific integrity committee that the results of
the DECREASE-IV trial are ``scientifically incorrect''
(p.11)\textsuperscript{3}.

@@

The effect of beta-blockade on perioperative mortality was already
unclear following the findings of scientific misconduct; these results
strongly affirm that the DECREASE trials should be neglected when
assessing the effectiveness of beta-blockers. Our results indicate that
the results from the DECREASE trials are nearly impossible to have
arisen from the same effect inspected by the non-DECREASE trials, except
when we assume at least some of the data were manipulated. As such, the
scientific validity of the DECREASE-I and DECREASE-IV trials should by
now be clear: they are to be regarded as irrelevant when assessing the
effectiveness of beta-blockade on perioperative mortality. Nonetheless,
the original papers that presented these trial results are not yet
retracted\textsuperscript{9,10}, something we strongly recommend based
on the combination of our results and the integrity
reports\textsuperscript{2--4}.

The ESC/ESA and ACC/AHA guidelines\textsuperscript{19,20} on
perioperative beta-blockade already excluded the DECREASE trials in
their assessment, but also state that other trials by Poldermans are
excluded. However, upon close inspection of the reference lists, the
ACC/AHA guidelines still cites four trials as evidence
base\textsuperscript{10,21--23}, of which two were already inspected by
the scientific integrity committees of Erasmus
MC\textsuperscript{10,21}. In the ACC/AHA guidelines, the following is
said about studies conducted by Poldermans:

\begin{quote}
\emph{``If nonretracted DECREASE publications and/or other derivative
studies by Poldermans are relevant to the topic, they can only be cited
in the text with a comment about the finding compared with the current
recommendation but should not form the basis of that recommendation or
be used as a reference for the recommendation.''}\textsuperscript{20}
\end{quote}

Nonetheless, references \emph{are} made without clear comments. Given
the confirmation of problems in the DECREASE-I and DECREASE-IV trials in
our results, it stresses that there is reason to distrust trials by
Poldermans. We pose that investigations should be initiated into works
where Poldermans was clearly involved and which were not cleared by the
scientific committees of Erasmus MC in their misconduct investigations.
Especially those papers cited as evidence in the ACC/AHA guidelines
should be investigated, considering they affect patients and their
treatment directly.

Previously, further investigation of trials by Poldermans was deemed
unfeasible due to the lack of raw data; here we indicate methods that do
make it feasible. Based on just event-count data and comparable trials,
we were able to estimate whether part of the data were in fact
fabricated and whether the results were within reason of comparable
trials. The results clearly indicated they were not.

The results of our analyses also highlight that, despite the lack of
availability of the raw data, summary results from larger samples allow
for better estimates of the number of fabricated data points when
similar trials are available. Moreover, larger trials result in
relatively more certainty (e.g., DECREASE-IV) about the estimated number
of fabricated data points, when using the inversion method, compared to
smaller trials (e.g., DECREASE-I). This increased certainty is due to
decreased standard errors at the population level of the estimated
effects, resulting in higher sensitivity to data anomalies. Nonetheless,
much residual uncertainty remains and simply fewer information is
available in the summary results. Raw data availability would improve
the options open to detect potential anomalies (note: raw data are
available for DECREASE VI, but upon a freedom of information request
Erasmus MC refused to share these data). The results also highlight that
in order to prevent detection, it would be in the fabricators' interest
to fabricate small studies even if raw data are hidden away (assuming
the fabricator wants to remain undetected).

With respect to clinical practice, the results provide some tentative
evidence that type of beta-blockade can severely influence perioperative
mortality. Our reanalysis of the Bouri et al.\textsuperscript{7} data
indicates that type of beta-blockade can reverse the effect on
perioperative mortality, even after taking into account whether a study
belongs to the DECREASE family. As such, atenolol seems to tentatively
decrease perioperative mortality, whereas the others (metoprolol,
propanolol, bisoprolol) increase perioperative mortality. However, there
seems to be covariation with respect to treatment administration,
duration, and dose, which further confounds whether the treatment effect
is due to type of beta-blocker or due to one of these other parameters.
This affirms the statement from the ESC/ESA guidelines that \emph{``high
priority needs to be given to new randomized clinical trials to better
identify which patients derive benefit from beta-blocker therapy in the
perioperative setting, and to determine the optimal method of
beta-blockade''}\textsuperscript{19}.

In sum, our research indicates that the DECREASE trials are nearly
impossible and contain at least some manipulated data points. We
recommend retraction of the DECREASE-I and DECREASE-IV trials, and
recommend renewed investigations into Poldermans' work --- especially
that work still referenced by guidelines on the use of beta-blockers.
Moreover, it remains unclear whether beta-blockers might be effective
given the right treatment. We recommend new and more extensively
controlled, confirmatory trials to determine whether there is any use in
administering beta-blockers in order to decrease perioperative mortality
--- at the moment there is insufficient evidence to determine any useful
effect of beta-blockers.

@@

\hypertarget{ref-dunkelgrun2009}{}
9. Dunkelgrun M, Boersma E, Schouten O, et al. Bisoprolol and
fluvastatin for the reduction of perioperative cardiac mortality and
myocardial infarction in intermediate-risk patients undergoing
noncardiovascular surgery: A randomized controlled trial (DECREASE-IV).
\emph{Annals of surgery}. 2009;249(6):921-926.
doi:\href{https://doi.org/10.1097/SLA.0b013e3181a77d00}{10.1097/SLA.0b013e3181a77d00}.

\hypertarget{ref-poldermans1999}{}
10. Poldermans D, Boersma E, Bax JJ, et al. The effect of bisoprolol on
perioperative mortality and myocardial infarction in High-Risk patients
undergoing vascular surgery. \emph{The New England journal of medicine}.
1999;341(24):1789-1794.
doi:\href{https://doi.org/10.1056/NEJM199912093412402}{10.1056/NEJM199912093412402}.

\hypertarget{ref-Buyse1999-jq}{}
11. Buyse M, George SL, Evans S, et al. The role of biostatistics in the
prevention, detection and treatment of fraud in clinical trials.
\emph{Statistics in medicine}. 1999;18(24):3435-3451.
doi:\href{https://doi.org/10.1002/(SICI)1097-0258(19991230)18:24\%3C3435::AID-SIM365\%3E3.0.CO;2-O}{10.1002/(SICI)1097-0258(19991230)18:24\textless{}3435::AID-SIM365\textgreater{}3.0.CO;2-O}.

\hypertarget{ref-Knepper2016-la}{}
12. Knepper D, Lindblad AS, Sharma G, et al. Statistical monitoring in
clinical trials: Best practices for detecting data anomalies suggestive
of fabrication or misconduct. \emph{Therapeutic Innovation \& Regulatory
Science}. 4\textasciitilde{}feb 2016.
doi:\href{https://doi.org/10.1177/2168479016630576}{10.1177/2168479016630576}.

\hypertarget{ref-klei2015}{}
13. Klei W van. Welke perioperatieve bètablokker heeft de voorkeur?
{[}Which perioperative beta-blocker is preferred?{]}. \emph{Nederlands
Tijdschrift voor Geneeskunde}. 2015;159:A9798.

@@

\hypertarget{ref-Kristensen_2014}{}
19. Kristensen SD, Knuuti J, Saraste A, et al. 2014 ESC/ESA guidelines
on non-cardiac surgery. \emph{European Journal of Anaesthesiology}.
2014;31(10):517-573.
doi:\href{https://doi.org/10.1097/eja.0000000000000150}{10.1097/eja.0000000000000150}.

\hypertarget{ref-Fleisher_2014}{}
20. Fleisher LA, Fleischmann KE, Auerbach AD, et al. 2014 ACC/AHA
guideline on perioperative cardiovascular evaluation and management of
patients undergoing noncardiac surgery: Executive summary: A report of
the american college of cardiology/american heart association task force
on practice guidelines. \emph{Circulation}. 2014;130(24):2215-2245.
doi:\href{https://doi.org/10.1161/cir.0000000000000105}{10.1161/cir.0000000000000105}.

\hypertarget{ref-Boersma_2001}{}
21. Boersma E, Poldermans D, Bax JJ, et al. Predictors of cardiac events
after major vascular surgery. \emph{JAMA}. 2001;285(14):1865.
doi:\href{https://doi.org/10.1001/jama.285.14.1865}{10.1001/jama.285.14.1865}.

\hypertarget{ref-van_Kuijk_2009}{}
22. Kuijk J-P van, Flu W-J, Schouten O, et al. Timing of noncardiac
surgery after coronary artery stenting with bare metal or drug-eluting
stents. \emph{The American Journal of Cardiology}.
2009;104(9):1229-1234.
doi:\href{https://doi.org/10.1016/j.amjcard.2009.06.038}{10.1016/j.amjcard.2009.06.038}.

\hypertarget{ref-Flu_2010}{}
23. Flu W-J, Kuijk J-P van, Chonchol M, et al. Timing of pre-operative
beta-blocker treatment in vascular surgery patients. \emph{Journal of
the American College of Cardiology}. 2010;56(23):1922-1929.
doi:\href{https://doi.org/10.1016/j.jacc.2010.05.056}{10.1016/j.jacc.2010.05.056}.
